---
output: 
  github_document:
    md_extensions: -autolink_bare_uris
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r packages, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(cache=TRUE)
if (!require("pacman")) install.packages("pacman")
p_load(data.table, feather, dplyr, tidyr, janitor, doParallel)
```

# OEC Yearly Datasets

## Description

Scripts to recreate the trade data available at the [Observatory of Economic Complexity](http://atlas.media.mit.edu/en/). 

We have a repo explaining the methodological details. Please visit [OEC Documentation](https://observatory-economic-complexity.github.io/oec-documentation/) and the [Official website](http://atlas.media.mit.edu/en/) for the details. 

These scripts are released under MIT license.

## How to use this project

While you can run the scripts from UNIX terminal, we highly recommend [RStudio](https://www.rstudio.com/).

Each project is organized as an RStudio project. Each project contains a master script titled `0-some-function.R` at the top level but some projects such as [OEC Product Space](https://github.com/observatory-economic-complexity/oec-product-space) are organized in subfolders where each subfolder contains a master script.

The master script, for any case, runs a function with user prompt and/or simple scripts. The safe way to run this without errors is to open the project and then run the master script from RStudio.

[OEC Packages Snapshot](https://github.com/observatory-economic-complexity/oec-packages-snapshot) stores all packages and its dependencies used across all repositories under this organization. To ease reproducibility [OEC Bundles](https://github.com/observatory-economic-complexity/oec-bundles) contains several bundles to avoid dependencies or packages related problems, and this project already contains an unbundled set of packages.

You can read more about RStudio projects [here](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects).

## Project structure

```
oec-yearly-datasets
└── README.md
└── README.Rmd
└── 1-raw-data
|    └── 0-download-function.R
|    └── hs-rev1992
|    └── hs-rev1996
|    └── hs-rev2002
|    └── hs-rev2007
|    └── hs-rev2012
|    └── sitc-rev2
└── 2-processed-data
|    └── 0-organize-function.R
|    └── hs-rev1992
|    └── hs-rev1996
|    └── hs-rev2002
|    └── hs-rev2007
|    └── hs-rev2012
|    └── sitc-rev2
└── packrat
```

## Output

The scripts generates different files organized under `1-raw-data/classification-rev-bulk/` (i.e `1-raw-data/hs-rev1992/`) and `2-processed-data/classification-rev/`.

### Generated files (HS 92 4 digits, 2015)

#### Raw data

```{r raw, eval=TRUE, echo=FALSE, message=FALSE}
system("7z e -aos 1-raw-data/hs-rev1992/zip/type-C_r-ALL_ps-2015_freq-A_px-H0_pub-20171207_fmt-csv_ex-20171228.zip")
skimr::skim(
  fread("type-C_r-ALL_ps-2015_freq-A_px-H0_pub-20171207_fmt-csv_ex-20171228.csv",
        colClasses = c("Commodity Code" = "character",
                       "Qty" = "numeric",
                       "Netweight (kg)" = "numeric",
                       "Trade Value (US$)" = "numeric"),
        showProgress = F)
) %>% skimr::kable()
```

#### Verified data

```{r verified, eval=TRUE, echo=FALSE, message=FALSE}
system("7z e -aos 2-processed-data/hs-rev1992/hs-rev1992-2015.zip")
skimr::skim(
  read_feather("hs-rev1992-2015.feather")
) %>% skimr::kable()
```

For each NA or 0 import/export value we tried to fill the gap. If country A reported NA or 0 exports (imports) of product B to (from) country C, then we searched what country C reported of imports (exports) of product B from (to) country A. The column `marker` indicated those replacements under this labels:

```
|marker|meaning                              |
|------|-------------------------------------|
|1     |imports with replacements            |
|2     |exports with replacements            |
|3     |imports and exports with replacements|
|NA    |no replacements needed               |
```

```{r rem, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
try(file.remove("type-C_r-ALL_ps-2015_freq-A_px-H0_pub-20171207_fmt-csv_ex-20171228.csv"), silent = TRUE)
try(file.remove("hs-rev1992-2015.feather"), silent = TRUE)
```


## Software details

Here the version information about R, the OS and attached or loaded packages for this project:

```{r sessioninfo, eval=TRUE, echo=FALSE}
sessionInfo()
```

## The MIT License

Copyright (c) 2017, Datawheel

Permission is hereby granted, free of charge, to any person obtaining
a copy of this software and associated documentation files (the
"Software"), to deal in the Software without restriction, including
without limitation the rights to use, copy, modify, merge, publish,
distribute, sublicense, and/or sell copies of the Software, and to
permit persons to whom the Software is furnished to do so, subject to
the following conditions:

The above copyright notice and this permission notice shall be
included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND
NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE
LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION
OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION
WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.

